### 2021-03-10 latex/questions.tex

\documentclass{article}
\usepackage{physics,amsmath}
\begin{document}

Page 181~\cite{giunti}: It is also important to note that the presence of sterile right-handed neutrino
fields is totally irrelevant for the cancellation of quantum anomalies [31, 201], which
constrain the properties of the other elementary fermion fields [268, 572, 517, 516,
606, 805, 915, 551, 459, 692]. As a consequence, the number of sterile right-handed
neutrino fields is not constrained by the theory, and the introduction of three right-
handed neutrino fields, one for each generation, is not even the minimal extension
of the SM, because the presence of only one right-handed neutrino field cannot be
excluded.

What are the quantum anomalies referring to?

The it is impossible to find a transformation that leaves both the diagonalized Higgs-lepton Yukawa Lagrangian and the right-handed neutrino fields invariant under global U(1) symmetry. So flavour lepton number is broken by sterile neutrinos? Obviously flavour is not conserved, but I thought that this was due to the introduction of mass, not the introduction of right-handed neutrinos.
Ans: If we only have Dirac mass, we make a minimal extension of SM. By introducing more than 1 neutrinos, and a RH- one, 6.55 becomes a matrix eq, and flavour oscilaltes. Wheter the matrix is diagonal or not is dictated by the data.
we have two U(1) symmetris: total lepton number (no flavor). 6.63 breaks this by 2 units (net u1 charge of 2 instead of 0). Outside of osc, flavor is conserved.

CP and T violations depend on the Dirac phases, and if neutrino experiments are independent of the Majorana phases, what is their significance?
Ans: If neutrinos are Dirac, we would only have one Dirac phase (as in quarks). To observe

"We will show that these flavor states reduce to the standard ones in eqn (7.4) in the case of experiments which are not sensitive to the
differences of the contributions of the different neutrino masses to the production
and detection processes. This is the case for all neutrino oscillation experiments."
Is this "differences of the contributions" referring to $\Delta m_{kj}^2$ or just $m_{kj}^2$?

Can IceCube detect CP asymmetry? I.e. is it sensitive to the oscillations of the flavor transition probabilities.

Next month 
Write code for 3 generation neutrino oscillation probability and extend to 3+x, x=0,1,2. 
Code for matter density through earth. Runge kutta is slow. 
Write own code for icecube

I will do both NC and CC. Data classified as cascade events (from NC and CC-electron/tau), or muon events (CC-muon).
NC: neutrino + nucleon(quarks) + Z = neutrino + other nucleons (t-channel) (can be DIS)
hadrons have strong interactions so they shower like crazy

CC neutrino + quarks +W = charged lepton (easy to see)
electron creates shower through bremsstrahlung. muon doesn't

Which parameters can I assume are given, and which parameters will be my "output"?
Should I focus/read up more on atmospheric or solar neutrinos?

Is there a reason to doubt unitarity of the PMNS matrix? Do sterile neutrinos break unitarity?

Why does large mass-squared splittings with sterile neutrinos imply non-unitarity?

P 169 in \cite{heros2020}: Sterile neutrinos produce a very fast oscillations at the ow-energy part of the atmospheric neutrino flux. This that Deepcore may not be sensitive to the sterile neutrinos.

What is the purpose of forecasting data? How is it done, and why?
\bibliographystyle{apalike}
\bibliography{notes}
\end{document}